The Chemistry Development Kit (CDK) v2.0: atom typing, depiction, molecular formulas, and substructure searching

### 2017-03-25 article.tex
Add ORCID
Small language corrections (rev #4)

\author[
   addressref={um},                                 % ORCID: 0000-0001-7542-0286
   %corref={aff1},
   %noteref={n1},
   email={egon.willighagen@maastrichtuniversity.nl}
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={nm},
   email={john.mayfield@cantab.net}
]{\inits{JWM}\fnm{John W} \snm{Mayfield}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={azg},                                % ORCID: 0000-0001-9491-4134
   email={Lars.A.Carlsson@astrazeneca.com}
]{\inits{LC}\fnm{Lars}~\snm{Carlsson}}
% \author[
%    addressref={jena},
%    email={kai.duehrkop@uni-jena.de}
% ]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina}~\snm{Jeliazkova}}
\author[
   addressref={leicester},
   email={shk12@le.ac.uk}
]{\inits{SK}\fnm{Stefan} \snm{Kuhn}}
\author[
   addressref={wi_mit},                             % ORCID: 0000-0002-6940-3006
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={miquel},                             % ORCID: 0000-0002-4659-1446
   email={mrojas@qca.es}
]{\inits{MRJ}\fnm{Miquel}~\snm{Rojas-Chertó}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={gilleain},                           % ORCID: 0000-0002-8368-6954
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris~T}~\snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},
  \postcode{NL-6200 MD},
  \city{Maastricht},
  \cny{The Netherlands}
}
\address[id=nm]{
  \orgname{NextMove Software Ltd},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{CB4 0EY},
  \city{Cambridge},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=azg]{
  \orgname{AstraZeneca, Innovative Medicines \& Early Development, Quantitative Biology},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Mölndal},
  \cny{SE}
}
% \address[id=jena]{
%   \orgname{Chair for Bioinformatics, Friedrich Schiller University},
%   %\street{},
%   \postcode{07743},
%   \city{Jena},
%   \cny{Germany}
% }
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=leicester]{
  \orgname{Department of Informatics, University of Leicester},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Leicester},
  \cny{UK}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{9 Cambridge Center},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=miquel]{
  \orgname{Química Clínica Aplicada},
  %\street{4 Hanway Place},
  \postcode{43870},
  \city{Amposta},
  \cny{Spain}
}
\address[id=gilleain]{
  %\orgname{Department of Informatics, University of Leicester},
  \street{4 Hanway Place},
  \postcode{W1T 1HD},
  \city{London},
  \cny{UK}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Metabolomics and Molecular Informatics, EMBL-EBI},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}

@@

  A reference implementation, \texttt{CDKAtomTypeMatcher}, has been
  written that perceives these atom types, and validates the
  perception automatically against the properties defined by the
  ontology.  This class handles a variety in types of missing
  information, as commonly resulting from various (file) formats; for
  example, it can handle undefined hydrogen counts and undefined
  double bond positions if hybridization information is provided
  instead.  That makes the perception code flexible but also more
  complex. Alternative algorithms for atom typing have not been
  explored. This reference implementation can be used on a single
  atom:

@@

  The perception from coordinates and wedge assignment algorithms are fundamental for conversion
  between formats that store stereochemistry implicitly based on coordinates (e.g. molfile, CML) and 
  explicitly (e.g. SMILES, CML, InChI). Perception from 2D coordinates can optionally identify
  perspective projects, specifically: Fischer, Haworth, and Chair projections. With the perception of 
  perspective projections enabled, database entries currently consider distinct can be merged
  (Figure~\ref{fig:stereoprojections}).  The perception is based on an  algorithm briefly  
  described by~\cite{Karapetyan2015}.

@@

  Graph matching of stereochemistry with the described representation is straight forward. Given
  the atom-atom mapping from a query structure to a target molecule, the focus and carriers of 
  the query stereochemistry are mapped to the target. Using the permutation parity of this mapping
  the configurations compared. SMARTS matching requires some special handling for complex cases 
  \cite{May2014_SMARTS}. For canonicalization, a partial canonical ordering is used to assign an 
  absolute label which can then be integrated into the ordering. More detailed descriptions of 
  these algorithms are described in detail in Chapter~6 of~\cite{May2015}.

>>>>

\author[
   addressref={um},                                 % ORCID: 0000-0001-7542-0286
   %corref={aff1},
   %noteref={n1},
   email={egon.willighagen@maastrichtuniversity.nl}
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={nm},                                 % ORCID: 0000-0001-7730-2646
   email={john.mayfield@cantab.net}
]{\inits{JWM}\fnm{John W} \snm{Mayfield}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={azg},                                % ORCID: 0000-0001-9491-4134
   email={Lars.A.Carlsson@astrazeneca.com}
]{\inits{LC}\fnm{Lars}~\snm{Carlsson}}
% \author[
%    addressref={jena},
%    email={kai.duehrkop@uni-jena.de}
% ]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina}~\snm{Jeliazkova}}
\author[
   addressref={leicester},
   email={shk12@le.ac.uk}
]{\inits{SK}\fnm{Stefan} \snm{Kuhn}}
\author[
   addressref={wi_mit},                             % ORCID: 0000-0002-6940-3006
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={miquel},                             % ORCID: 0000-0002-4659-1446
   email={mrojas@qca.es}
]{\inits{MRJ}\fnm{Miquel}~\snm{Rojas-Chertó}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={gilleain},                           % ORCID: 0000-0002-8368-6954
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris~T}~\snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},
  \postcode{NL-6200 MD},
  \city{Maastricht},
  \cny{The Netherlands}
}
\address[id=nm]{
  \orgname{NextMove Software Ltd},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{CB4 0EY},
  \city{Cambridge},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=azg]{
  \orgname{AstraZeneca, Innovative Medicines \& Early Development, Quantitative Biology},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Mölndal},
  \cny{SE}
}
% \address[id=jena]{
%   \orgname{Chair for Bioinformatics, Friedrich Schiller University},
%   %\street{},
%   \postcode{07743},
%   \city{Jena},
%   \cny{Germany}
% }
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=leicester]{
  \orgname{Department of Informatics, University of Leicester},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Leicester},
  \cny{UK}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{455 Main Street},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=miquel]{
  \orgname{Química Clínica Aplicada},
  %\street{4 Hanway Place},
  \postcode{43870},
  \city{Amposta},
  \cny{Spain}
}
\address[id=gilleain]{
  %\orgname{Department of Informatics, University of Leicester},
  \street{4 Hanway Place},
  \postcode{W1T 1HD},
  \city{London},
  \cny{UK}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Metabolomics and Molecular Informatics, EMBL-EBI},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}

@@

  A reference implementation, \texttt{CDKAtomTypeMatcher}, has been
  written that perceives these atom types, and validates the
  perception automatically against the properties defined by the
  ontology.  This class handles a variety of types of missing
  information, as commonly resulting from various (file) formats; for
  example, it can handle undefined hydrogen counts and undefined
  double bond positions if hybridization information is provided
  instead.  That makes the perception code flexible but also more
  complex. Alternative algorithms for atom typing have not been
  explored. This reference implementation can be used on a single
  atom:

@@

  The perception from coordinates and wedge assignment algorithms are fundamental for conversion
  between formats that store stereochemistry implicitly based on coordinates (e.g. molfile, CML) and 
  explicitly (e.g. SMILES, CML, InChI). Perception from 2D coordinates can optionally identify
  perspective projections, specifically: Fischer, Haworth, and Chair projections. With the perception of 
  perspective projections enabled, database entries currently consider distinct can be merged
  (Figure~\ref{fig:stereoprojections}).  The perception is based on an  algorithm briefly  
  described by~\cite{Karapetyan2015}.

@@

  Graph matching of stereochemistry with the described representation is straight forward. Given
  the atom-atom mapping from a query structure to a target molecule, the focus and carriers of 
  the query stereochemistry are mapped to the target. Using the permutation parity of this mapping
  the configurations were compared. SMARTS matching requires some special handling for complex cases 
  \cite{May2014_SMARTS}. For canonicalization, a partial canonical ordering is used to assign an 
  absolute label which can then be integrated into the ordering. More detailed descriptions of 
  these algorithms are described in detail in Chapter~6 of~\cite{May2015}.
